---
output: md_document
---
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
  version.github <- as.vector(read.dcf('DESCRIPTION')[, 'Version'])
  version.github <- gsub('-', '.', version.github)

  # get the version from the CRAN package mirror
  cran.dcf.con <- url("https://raw.githubusercontent.com/cran/lodi/master/DESCRIPTION")
  version.cran <- read.dcf(file = cran.dcf.con)[, "Version"]
  close(cran.dcf.con)
  if (version.cran == version.github) {
      color <- "success"
  } else {
      color <- "informational"
  }


```
<!-- badges: start -->
[![CRAN Version](https://www.r-pkg.org/badges/version/lodi)](https://cran.r-project.org/package=lodi)
[![Github Version](https://img.shields.io/badge/Github-`r version.github`-`r color`.svg?style=flat)](https://github.com/umich-cphds/lodi)
[![Travis CI](https://travis-ci.org/umich-cphds/lodi.svg?branch=master)](https://travis-ci.org/umich-cphds/lodi)
<!-- badges: end -->


# Limit of Detection Multiple Imputation

`lodi` is a R package that implements censored likelihood multiple
imputation (CLMI) for single pollutant models with exposure
biomarkers below their respective detection limits. `lodi` also
contains implementations for standard methods such as single
imputation with a constant and complete-case analysis, although
those methods are primarily designed for comparison with `clmi`.

## Installation
`lodi` requires `rlang >= 0.3.0` to be installed, so you may
want to install or update `rlang` before installing `lodi`.

The package can be installed from CRAN
```{r, eval = F}
install.packages("lodi")
```
Or from Github
```{r, eval = F}
# install.packages("devtools")
devtools::install_github("umich-cphds/lodi", build_opts = c())
```
The github version may contain bug fixes not yet present on CRAN,
so if you are experiencing issues, you may want to try the Github
version of the package.

## Example
Once `lodi` is installed, you can load up R and type

```
vignete("lodi")
```
to learn how to use the method.

## Bugs
If you encounter a bug, please open an issue on the [Issues](https://github.com/umich-cphds/lodi/issues) tab on Github or send
us an email.

## Contact
For questions or feedback, please email Jonathan Boss at
<bossjona@umich.edu> or Alexander Rix <alexrix@umich.edu>.

## References
Boss J, Mukherjee B, Ferguson KK, et al. Estimating outcome-exposure
associations when exposure biomarker detection limits vary across batches.
*Epidemiology*. 2019;30(5):746-755.
[10.1097/EDE.0000000000001052](https://doi.org/10.1097/EDE.0000000000001052)
