---
title: "Censored Likelihood Multiple Imputation in R"
author: "Jonathan Boss"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Censored Likelihood Multiple Imputation in R}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Loading `clmi` and an example dataset

For convenience we have included a example dataset called `toy.data`, which can
be loaded by running `data("toy-example")`. Let's look at the first 10 entries
of the example dataset.

```{r}
library(clmi)

data("toy-example")

head(toy.data, n = 10)
```

The `id` column gives the study ID and is unimportant for the purposes of this
example. The `case_cntrl` takes values 0 or 1, where 1 indicates that the
subject has the disease of interest and 0 indicates that the subject is a
healthy control. The column `poll` is the environmental exposure of interest,
where `NA` indicates that the concentration is below the limit of detection
(LOD). The variables `smoking` and `gender` are covariates that we are going to
include in the imputation model. Lastly, `batch1` takes two values; 1 if the
subject's biosample was assayed in batch 1 and 0 if the subject's biosample was
assayed in batch 2. Moreover, the LOD for batch 1 is 0.8 and the LOD for
batch 2 is 0.65.

## Implementing Censored Likelihood Multiple Imputation

The function that performs censored likelihood multiple imputation is the
`clmi` function. For details see `help(clmi)`.

```{r}
clmi.out <- clmi(df = toy.data, contaminant = "poll", batch = "batch1",
                 outcome = "case_cntrl", contam.covars = c("smoking", "gender"),
                 lod.info = data.frame(batch.info = c("1","0"),
                 lod = c(0.8, 0.65)), n.imps = 20, seed = 12345,
                 t.function = function(x) log(x))
```

Some important things to notice:

- The `contaminant`, `batch`, `outcome`, and `contam.covars` arguments must
  correspond to the variable names in `toy.data`.

- The `lod.info` argument is where all of the LOD information goes. The first
  column must be named `batch.info` and contains all levels of the `batch1`
  variable as character strings. The second column must be named `lod` and
  contains the LODs corresponding to each batch. Remember that all of the
  values below the LOD must be entered as `NA` so that the function recognizes
  them as missing.

- The `t.function` argument specifies whether you want to return an imputed
  concentrations on the untransformed scale (default) or on a transformed scale.
  It is often the case that contaminant concentrations are log-transformed to
  handle right skewed exposure data while still maintaining interpretable effect
  estimates. To impute on the log-scale specify,
  `t.function = function(x) log(x)`. As another example, if you want to impute
  on the square root scale then specify, `t.function = function(x) sqrt(x)`.

The imputed datasets can be extracted as a list using `$imputed.dfs`:

```{r, eval=FALSE}
extract.imputed.dfs <- clmi.out$imputed.dfs
```

## Fit and pool outcomes models

The `pool.clmi` function takes the output generated by the `clmi` function, fits
outcome models on each of the imputed datasets, and pools inference across
outcome models. For details see `help(pool.clmi)`.

```{r}
results <- pool.clmi(clmi.out = clmi.out, regression.type = "logistic",
                     outcome.covars = NULL)
```

A few things that deserve further comment:

- There are currently two options for the `regression.type` argument. If you
  have binary outcome data (as in the current example) use
  `regression.type = "logistic"` so that the outcome models fit on the imputed
  datasets are logistic regression models. If you have continuous outcome data
  use `regression.type = "linear"` so that the outcome models fit on the
  imputed datasets are linear regression models.

- The `outcome.covars` argument is a vector of strings containing the variable
  names corresponding to the precision variables that you want to include as
  adjustment covariates in the outcome models. Note that the strings must
  correspond to variable names in your dataset. Variables included in the
  `contam.covars` when running the `clmi` function are automatically adjusted
  for. Therefore, `outcome.covars` should only contain additional adjustment
  covariates that were not included in the imputation model. In this case
  `outcome.covars = NULL` means that we only want an single-pollutant outcome
  model that adjusts for the variables in `contam.covars`, i.e. `smoking` and
  `gender`.

To display the pooled results use `$output`:

```{r}
results$output
```

If you want to look at the individual regressions fit on each imputed dataset use
  `$regression.summaries`

```{r, eval=FALSE}
results$regression.summaries
```

##Contact information

If you would like to report a bug in the code, ask questions, or send
requests/suggestions e-mail Jonathan Boss at `bossjona@umich.edu`.
